Deployable Probabilistic Programming

### 2019-08-11 paper/paper.tex
intro example of prob prog

@@

@@

The eight schools problem~\cite{GCS+03} is an application of
the hierarchical normal model that often serves as an introductory
example of Bayesian statistical modeling. Without going into
details, in this problem the effect of coaching programs on test
outcomes is compared among eight schools.
Figure~\ref{fig:eight-schools} shows specifications of the same
model in Stan and Infergo, side-by-side.
\begin{figure*}
\begin{minipage}[t]{0.45\textwidth}
  \begin{lstlisting}[language=Stan,framexleftmargin=10pt,numbers=left]
data {
  int<lower=0> J;
  vector[J] y;
  vector<lower=0>[J] sigma;
}

@@

The model in Figure~\ref{fig:linear-regression} specifies linear
regression on unidimensional data: an improper uniform prior is
placed on the parameters $\alpha$, $\beta$, and $\log \sigma$
(lines 9--10), which are then conditioned on observations $x, y$
such that
\begin{equation}
y \sim \mathrm{Normal}(\alpha + \beta x, \sigma)
	\label{eqn:linear-regression-1}
\end{equation}
(lines 12--17).
\begin{figure}
\begin{lstlisting}[framexleftmargin=10pt,numbers=left]
type LRModel struct {
  Data  [][]float64
}

@@

The Gaussian mixture model is a probabilistic model of a mixture
of components, where each component comes from a Gaussian
distribution. The model variant in
Figure~\ref{fig:gaussian-mixture} accepts a vector of
single-dimensional observations and the number of components,
and infers the distribution of parameters of each component
in the mixture. Following Stan User's Guide~\cite{SDT18},
component memberships are summed out to make the model
differentiable.
\begin{figure}
\begin{lstlisting}[framexleftmargin=10pt,numbers=left]
type GMModel struct {
  Data  []float64 // observations
  NComp int       // number of components
}

>>>>

\lstdefinelanguage{WebPPL}{
  keywords={typeof, new, true, false, catch, function, return,
  null, catch, switch, var, if, in, while, do, else, case,
  break, condition},
  ndkeywords={class, export, boolean, throw, implements, import, this},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]"
}

@@

\begin{figure}
\begin{lstlisting}[language=WebPPL]
var breastCancer = flip(0.01)
var benignCyst = flip(0.2)
var positiveMammogram = (breastCancer && flip(0.8))
                      || (benignCyst && flip(0.5))
condition(positiveMammogram)
return breastCancer
\end{lstlisting}
	\caption{A probabilistic program in WebPPL~\cite{GS15}:
	probability of breast cancer given a positive mammogram.}
	\label{fig:mammo}
\end{figure}
  Consider, for example, the probabilistic program in
  Figure~\ref{fig:mammo}. The program models dependency between
  a positive mammogram and possible causes --- breast cancer
  or a benign cyst.  The program defines three Bernoulli
  random variables  --- \lstinline{breastCancer},
  \lstinline{benignCyst}, \lstinline{positiveMammogram} ---
  and conditions on one of the random variables,
  \lstinline{positiveMammogram}, being true. Inference on the
  program returns the probability of breast cancer given a
  positive mammogram.

@@

The eight schools problem~\cite{GCS+03} is an application of
the hierarchical normal model that often serves as an introductory
example of Bayesian statistical modeling.  In this problem,
experiments were performed to estimate the effects of coaching
programs in each of eight high schools. The outcome variable in
each study was the increase in test score compared to the same
test before the program. The average increase and the standard
deviation of the increase were reported for each school. Based
on that and on assumption of similarity between schools,
expected effects are to be estimated.
Figure~\ref{fig:eight-schools} shows specifications of the same
model in Stan and Infergo, side-by-side.
\begin{figure*}
\begin{minipage}[t]{0.45\textwidth}
  \begin{lstlisting}[language=Stan,framexleftmargin=10pt,numbers=left]
data {
  int<lower=0> J;
  vector[J] y;
  vector<lower=0>[J] sigma;
}

@@

The model in Figure~\ref{fig:linear-regression} specifies linear
regression (approximating linear dependency between input $x$
and output $y$) on unidimensional data. An improper uniform
prior is placed on the parameters $\alpha$,
$\beta$, and $\log \sigma$ (lines 9--10), which are then
conditioned on observations $x, y$ such that
\begin{equation}
y \sim \mathrm{Normal}(\alpha + \beta x, \sigma)
	\label{eqn:linear-regression-1}
\end{equation}
(lines 12--17).
\begin{figure}
\begin{lstlisting}[framexleftmargin=10pt,numbers=left]
type LRModel struct {
  Data  [][]float64
}

@@

The Gaussian mixture model is a probabilistic model of a mixture
of components, where each component comes from a Gaussian
distribution. Mixture models are used to represent a population
consisting of different sub-populations, without indication in
the data to which sub-population each observation belongs. The
model variant in Figure~\ref{fig:gaussian-mixture} accepts a
vector of single-dimensional observations and the number of
components, and infers the distribution of parameters of each
component in the mixture. Following Stan User's
Guide~\cite{SDT18}, component memberships are summed out to make
the model differentiable.
\begin{figure}
\begin{lstlisting}[framexleftmargin=10pt,numbers=left]
type GMModel struct {
  Data  []float64 // observations
  NComp int       // number of components
}
